---
title: "Business Analytics - ETC3250 2018 - Lab 8"
author: "Souhaib Ben Taieb"
date: "19 April 2018"
output: pdf_document
subtitle: The bootstrap
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "#",
  fig.height = 4,
  fig.width = 8,
  fig.align = "center",
  cache = FALSE
)
```

### Exercise 1

Do the exercise 2 in Section 5.4 of ISLR.




### Bootstrap confidence interval of the correlation coefficient

We will find a 95% confidence interval for the correlation coefficient of Median House value and average number of rooms in the Boston data set from the `MASS` package.

```{r, echo=FALSE, message=FALSE}
library(ISLR)
library(ggplot2)
```

The functions `cor` and `cor.test` will compute the correlation and an asymptotic 95% confidence interval for it. This interval is based on Fisher's z transform
$$ z = \frac{1}{2}\log\left(\frac{1+r}{1-r}\right)$$
which is approximately normally distributed with variance $1/(n-3)$ where $n$ is the number of observations. So if $z_L$ and $z_U$ are upper and lower limits for $z$, then 
$$
r_L = \frac{\exp(2z_L) - 1}{\exp(2z_L)+1}
  \qquad\text{and}\qquad 
r_U = \frac{\exp(2z_U) - 1}{\exp(2z_U)+1}
$$
are upper and lower limits for $r$.

We will use the bootstrap to test if this is a good approximation in this case.

### Exercise 2

Check that the confidence interval returned by `cor.test` is computed using the above transformation. 


### Exercise 3

Compute a 95% bootstrap confidence interval for the correlation. You will need to sample rows of the `Boston` matrix.


###  Exercise 4

Write a function that will return a bootstrap confidence interval for the correlation of any two numeric variables of the same length. Your function should take four arguments: 

 - `x`: a numeric vector of data
 - `y`: a numeric vector of data
 - `level`: the probability coverage of the confidence interval with default value of 0.95 
 - `B`: the number of bootstrap samples with default value of 1000.
